Enhanced Particle Swarm Optimization In Sketch Recognition

### 2010-08-16 trunk/Swarm/tmp/swarmModified.tex
correcting

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm. The enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols. Results show that that the enhanced particle swarm optimization improves the final recognition accuracy. 
 % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
%%% old abstract.. 
 %Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture. This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm. Simulation results will be carried out on stock market indices as an example of time series prediction which shows that our proposed training algorithm is more efficient and accurate than the Basic-PSO in stock market prediction.
\end{abstract}
\pagenumbering{arabic}
 \section{Introduction}

@@

This paper modifies the sketch recognition system presented in \cite{mypaper}, to use Enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes. As in the Orginal system the users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different EPSO algorithms. Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}. To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced. The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.

The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.  Enhanced Particle Swarm Algorithm is discribed in Section \ref{sec:EnhancedSwarmAlgorithm}. Section  \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process. Section \ref{sec:Experiments} describes the experiments that were preformed. We conclude in Section \ref{ConclusionandFutureWork}.

@@

\section{Enhanced Particle Swarm}
\label{sec:EnhancedSwarmAlgorithm}
In the above basic PSO, the velocity update is based on three terms an interia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term. The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions. Therefore, we are tyring to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term. The modified term adds social behvaiour of other particle and keep diversity while searching for the global optimum. The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.

@@

%where $iRandom_i$ and $Random_j$ are two particles in the solution space. This modification replace the distance between the current particle and the local best with two random particle from the space. The effect of this change in more expolration and avoiding of local min problems. Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated. The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used. 
\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recoginition system in \cite{myPaper}. The 
%\subsection{The system}

 The suggested sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

@@

\subsection{Non ellipse fitting algorithm}
\label{subsubsec:Discreteparticleswarmalgorithm}
\begin{figure}
	\centering
		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{ppd.jpg}}
	%	\hfill
	 \includegraphics[scale=0.7]{pso1.jpg}			
	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}
Two DPSO algorithms are used to generate different segmentations for each stroke the user draws. For each segmentation generated an error is evaluated. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. 
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

@@

After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

@@

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectivily. The results show that the enhanced swarm optimization algorithm improve the segmentation of the system which leds to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistant in all swarm algorithms and in both experiments.    %n the We performed two experiments to test the system. Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization. To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.

@@

This paper used an  Enhanced-PSO  to modify a new sketch recognition using DPSO represented in \cite{mypaper}. Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The final recognition improved by modfiying the particle optimization algorithm without depending on the higher level features. The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset. Results show that the enhanced PSO algorithm improves the segmenattion on both algorithms which achieves better segmentation and final accuracy. The results proved that the Enhancment made on the PSO improves the efficency of the Basic PSO algorithm without being dependable on special objective function or application. The system achieved an average overall improvement of more than 1\% over Basic PSO.

 For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm. Also, a possible extentsion of this research is to improve the recognition algorithm of the Sketch Recognition system. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.   %this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification

>>>>

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm. The enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols. Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.

 % sketch are divided Users draw symbols and sketches using a set of pen strokes. 
%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
%%% old abstract.. 
 %Also another enhanced Particle Swarm Optimization (EPSO) based on the perturbation of the particle velocity is introduced to train the weights of this proposed neural network architecture. This proposed architecture with its relevant training algorithm is mentioned as (SDRNN-EPSO) and its performance is compare with the proposed network architecture trained be Basic-PSO algorithm. Simulation results will be carried out on stock market indices as an example of time series prediction which shows that our proposed training algorithm is more efficient and accurate than the Basic-PSO in stock market prediction.
\end{abstract}
\pagenumbering{arabic}
 \section{Introduction}

@@

This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes. As in the Original system the users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different EPSO algorithms. Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}. To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced. The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.

The remaining of the paper is as following; Section \ref{sec:ParticleSwarmAlgorithm} describes the Basic particle swarm algorithm in details.  Enhanced Particle Swarm Algorithm is described in Section \ref{sec:EnhancedSwarmAlgorithm}. Section \ref{Sysdisc} provide details on preprocessing, segmentation and recognition process. Section \ref{sec:Experiments} describes the experiments that were preformed. We conclude in Section \ref{ConclusionandFutureWork}.

@@

\section{Enhanced Particle Swarm}
\label{sec:EnhancedSwarmAlgorithm}
In the above basic PSO, the velocity update is based on three terms an inertia term, local cognitive ($l_{best}$) term and global cognitive ($g_{best}$)term. The particle local cognitive term is associated with the particle own history without consideration of other swarm particles locations and histories which leads to a lake of diversity and sub optimum solutions. Therefore, we are trying to add more social influence in the particle movement by using the weighted difference of the position vectors of any other two distinct particles randomly chosen from the swarm to modify the local cognitive term. The modified term adds social behavior of other particle and keeps diversity while searching for the global optimum. The new social cognitive term of the particle is shifted to the new location only if it yields a fitness value better than obtained using the standard PSO\cite{stockPaper}.

@@

%where $iRandom_i$ and $Random_j$ are two particles in the solution space. This modification replace the distance between the current particle and the local best with two random particle from the space. The effect of this change in more expolration and avoiding of local min problems. Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated. The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used. 
\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{myPaper}. 
%\subsection{The system}

 The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

@@

\subsection{Non ellipse fitting algorithm}
\label{subsubsec:Discreteparticleswarmalgorithm}
\begin{figure}
	\centering
		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{ppd.jpg}}
	%	\hfill
	 \includegraphics[scale=0.7]{pso1.jpg}			
	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}
Two DPSO algorithms are used to generate different segmentations for each stroke the user draws. For each segmentation generated, an error is evaluated. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. 
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

@@

After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

@@

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.    %n the We performed two experiments to test the system. Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization. To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.

@@

This paper used an Enhanced-PSO  to modify a new sketch recognition using DPSO represented in \cite{mypaper}. Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features. The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset. Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy. The results proved that the Enhancement made on the PSO improves the efficiency of the Basic PSO algorithm without being dependable on special objective function or application. The system achieved an average overall improvement of more than 1\% over Basic PSO.

 For the future work, it is suggested that we continue to improve both the enhanced PSO segmentation algorithm. Also, a possible extension of this research is to improve the recognition algorithm of the Sketch Recognition system. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.   %this new enhanced architecture will be used in other different applications of time series prediction and also extended to be used in other applications rather than time series prediction such as pattern recognition and speech verification
